---
output: html_document
---
#HW 2 - La Quinta is Spanish for Next to Denny's#
Team 4 - Duke Dinosaurs    
Due: Oct 14, 2015  

##LaQuinta##
There are two steps to scraping all the LaQuintas' across the United States. First we need a `get_lq.R` file which downloads all the htmls of the LaQuintas across the United States. Then we need a `parse_lq.R` file which searches the individual htmls for the appropriate information. 

On the most rudimentary level, the `get_lq.R` file uses two commands to download the htmls. First, we need to specify a `base_url`. The `base_url` is the top-most level-- "www.lq.com" -- that R needs to know. When we define this `base_url`, R will access the specified url to execute the next command. The next line, `listings_page`, will go to a secondary level from the `base_url`. Thus, in our case, R will go to this page: "http://www.lq.com/en/findandbook/hotel-listings.html." From here, we identified a .csv file that had the list of States with at least one LaQuinta, namely `lq_states.csv`. Then, in order to download the respective htmls of all the LaQuintas in each state, we utilized the function `get_state_hotels.` This function calls the `lq_states.csv` file and runs a loop for each State specified in the file thereby downloading all the corresponding htmls from the list of all the LaQuintas in the United States.  In order to prevent the LaQuinta homepage from blocking our download call, we implemented the "Sys.sleep(5)" command, which allowed for a 5 second lapse between each download call. Thus our `get_lq.R` function downloads all the htmls of the LaQuintas in the United States. 

Once we have downloaded all the htmls off of LaQuinta's webpage, we then want to parse through the htmls to compile the necessary information such as the address of the hotel, latitude and longitude, availability of internet, free parking, free complementary breakfast, swimming pools and etc. We were able to compile the information with by using selectorgadget on each of the individual hotel's pages. For example, if we want to know the above-mentioned information on LaQuinta's Durham-Chapel Hill location, we specified third level--the individual hotel's webpage--from the `listings_page`. Hence, to figure out the amenities in LaQuinta's Durham-Chapel Hill location, R will scrape the information from this page: "http://www.lq.com/en/findandbook/hotel-details.durham-chapel-hill.address.html". We used selectorgadget to pull the address, city, latitude, longitude and amenities. In order to construct a dataframe, we first specified a null dataframe to plug values into. Then we ran a loop to plug in the information that we scraped off of each hotel page. Then in the end, we used the rbind command to combine the rows into one dataframe `lq_final`.    


##Denny's##
To obtain all the Denny's locations in the US, we started by using a website called freemaptools to draw circles around parts of the US. With five circles, we should be covering all of US.
  
```
   Longitude Latitude Radius.Miles
1 -113.90625 40.64997          788
2  -95.44922 38.01140          851
3  -84.90234 38.88282         1056
4 -157.14844 20.82924          438
5 -141.67969 59.06586         1053
```
The first two columns tells us where the center of our circles are located. Radius.Miles tells us the radius of our circles in miles. These coordinates and radii will be stored in dennys_coords.csv so it can be used in the next steps.   
   
Now, in `get_dennys.R`, we'll extract the location data from `hosted.where2getit.com/dennys`. We get the locations in such a way because Denny's uses the third-party Where2GetIt API to display their restaurant locations. It will take the circle centers and radii specified in `dennys_coords.csv` to search for all locations within the radius of that circumference. We first specified the key, which Denny's uses, to access Where2GetIt API. Next, we limited the number of results to be 1,000. Now, `dennys_coords.csv` will have the longitude and latitude of the circle centers, as well as the search radius. Since we have five circles, we will run through the function five times using a ForLoop. The information will be stored as an .xml file under the dennys directory inside our data directory.    
   
Finally, we can construct a data frame with the details of each denny's location. We'll find the longitude, latitude, postal code, address, city, country, state, and unique id of each location. Since our circles overlap, we may have some locations listed multiple times. We'll use `unique` so that we only have each location listed once.    
Now that we have information on all Denny's locations in the US, it's possible to do distance analysis.    

##Distance Analysis##
    
We can see from our data that the number of Denny's is almost double that of LaQuintas. Intuitively, we might draw the conclusion that Denny's is more densely populated in the US. However, can we really say so? Is that description accurate? Now we do a little brainstorm with a simple example to illustrate the degree of clustering and separation in spatial data, and provide some evidence/disproof for Hedberg's claim.        
    
For our Distance Analysis, we first need to decide how we are going to calculate distance. We've decided to use the Haversine Formula. This formula is used to calculate the distances between two points on a sphere where the path will go along the surface of the sphere. Instead of the Euclidean distance, the formula will give us the great-circle distance. Although Earth is not a sphere, for our purposes, it is enough to approximate it to be a sphere with a radius of 3,959 miles.      
    
So the issue with finding the closest Denny's for a LaQuinta is that, that LaQuinta may not be the nearest LaQuinta to that Denny's. This creates a discrepancy. Instead, to conduct our distance analysis, we will calculate the distance to the nearest Denny's for all LaQuinta's and vice versa. We can then analyze our results to draw some conclusions about their locations.     
     
To illustrate we take an example where we have four LaQuintas and four Denny's. Let LaQuintas' coordinates be (1,0), (0,1), (1,1), and (0,0). Let Denny's coordinates be (0,0.001), (0.001,0), (-0.001,0), and (0,-0.001). So the closest LaQuinta to all the Denny's would be the one at (0,0). However, it wouldn't be true vice versa. In this example Denny's is more clustered and we can see that the average distance between a Denny's and the closest LaQuinta will be less than the average distance between a LaQuinta and the nearest Denny's.    
   
Denny's is a popular restaurant that will draw in any customer that would like to eat. On the other hand, LaQuinta is a hotel, so it will only draw in customers who are travelling and need a place to stay. As a result, LaQuintas will, most likely, be found in towns or cities that have a lot of visitors and tourists, which is why they will be clustered. 
     
In conclusion, our hypothesis is that LaQuinta will be more clustered than Denny's. In general, the distance between a Denny's to the closest LaQuinta will be greater than the distance between a LaQuinta and the closest Denny's.   

```{r}
load("data/dennys.Rdata") # loads dennys_df #
load("data/lq.Rdata") # loads lq_df #
library(magrittr)
spatial_dist <- function(thisx,thisy,thatx,thaty){
  # x is latitude, y is longitude #
  R = 3959 # Earth's radius in miles #
  radians = function(degrees) return(degrees*pi/180)
  # turns lat, long degrees into radians #
  this_lat = radians(thisx)
  this_long = radians(thisy)
  that_lat = radians(thatx)
  that_long = radians(thaty)
  # change lat and long to radians #
  a = (sin((that_lat - this_lat)/2))^2
  b = cos(this_lat)*cos(that_lat)*(sin((that_long - this_long)/2))^2
  d = 2*R*asin(sqrt(a+b)) # haversine formula #
  return(5280*d) # distance between this and that #
}

#The following function convert the coordinate data frame to a list, each element of the list a 
#two-tuple vector, containing latitude and longitude for each restaurant/hotel, we call this kind of data structure "clean list"
dataframe_tolist <- function(frame){
  split(frame, seq(nrow(frame)))
}
#The following part is a bit painful, trying to parse the two data frames for Dennys and lq to the "clean lists"
#The following two lines are quite straightfoward, extract the lat and long columns, and then convert the df to "clean lists" using
# the function just defined above
d_coord <- data.frame(latitude= dennys_df$Latitude,longitude=dennys_df$Longitude,stringsAsFactors = FALSE) %>% dataframe_tolist()
l_coord <- data.frame(latitude= lq_df$Latitude,longitude=lq_df$Longitude,stringsAsFactors = FALSE)  %>% dataframe_tolist()
#Then I find each vector has weird name on it so I make the names NULL
names(d_coord) <- NULL
names(l_coord) <- NULL
#Then I find each element in the "clean list" is actually a list instead of a vector, so I unlist each element
d_coord <- lapply(d_coord,unlist)
l_coord <- lapply(l_coord,unlist)
#Finally I find the content is each vector is character instead of numeric, so I make them double
d_coord <- lapply(d_coord,as.double)
l_coord <- lapply(l_coord,as.double)

#The following two functions are very very similar only excluding the second line.
#The reason why I write two similar functions twice is that I try to use lapply
#The input of distd function is each element of the "clean list" for dennys
#It returns a three-tuple, whose elements are this denny's latitude, longitude and the distance between this denny's and its nearest lq
distd <- function(this){
  that <- l_coord # make that the "clean list" of lq
  min_dist <- Inf # set min_dist as Infinity (so we can compare and find the min distance as below)
  for(i in that){ # for each hotel of lq
    temp_dist <- spatial_dist(unlist(this)[1],unlist(this)[2],unlist(i)[1],unlist(i)[2])
    # make temp_dist the patial distance between this dennys and that lq
    if(temp_dist < min_dist){
      min_dist <- temp_dist
      # if this number is smaller then min_dist, replace it, you know thus we can find the nearest distance
    }
  }
  return (c(unlist(this),min_dist))
  #return the three-tuple
}

#distl is sysmmtric
distl <- function(this){
  that <- d_coord
  min_dist <- Inf
  for(i in that){
    temp_dist <- spatial_dist(unlist(this)[1],unlist(this)[2],unlist(i)[1],unlist(i)[2])
    if(temp_dist < min_dist){
      min_dist <- temp_dist
    }
  }
  return (c(unlist(this),min_dist))
}

#now we apply distd on denny's "clean list"
dennys_dist <- lapply(d_coord, distd)
#Then we have made a dataframe with three columns, denny's lat, denny's long and dist_of_nearest_lq
dennys_neighbor <- do.call(rbind, dennys_dist)
colnames(dennys_neighbor) <- c("dennys_latitude","dennys_longitude","dist_of_nearest_lq")

#same for lq
lq_dist <- lapply(l_coord, distl)
lq_neighbor <-do.call(rbind, lq_dist)
colnames(lq_neighbor) <- c("lq_latitude","lq_longitude","dist_of_nearest_denny's")

#Basically we have obtained two sets of distances:
#From denny's point of view, the distance of the nearest lq
# and from lq's point of view, this distance of the nearest denny's
#below we plot the distributions for them
par(mfrow=c(1,2))
hist(dennys_neighbor[,3],breaks = 3000,xlim = c(0,2.0e+05))
hist(lq_neighbor[,3],breaks = 300,xlim = c(0,2.0e+05))

#if we zoom in a little, we can see it clearer
par(mfrow=c(1,2))
hist(dennys_neighbor[,3],breaks = 3000,xlim = c(0,2.0e+05))
hist(lq_neighbor[,3],breaks = 300,xlim = c(0,2.0e+05))


```

```{r}
summary(dennys_neighbor[,3])
summary(lq_neighbor[,3])
```

Finding some summary statistics about the distances between the nearest LaQuinta to all the Denny's and the nearest Denny's to all LaQuintas will tell us a little about how these two establishments are clustered in the US.    
These are our results. 
```
 summary(lq_neighbor[,3])
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    50.2   4369.0  16580.0  61850.0  62090.0 923900.0 
 summary(dennys_neighbor[,3])
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
      50    12780    33270   138500    92450 12690000 
  These measurements are in feet. 
```
As you can see, the summary statistic values for Denny's are greater than LaQuintas. This tells us that LaQuintas are more clustered than Denny's, which can be verified by going onto Google Maps and pulling up all the LaQuinta and Denny's locations in the US. You'll see that Denny's is significantly more widespread.    
   
This result isn't surprising and proves our earlier hypothesis.  
    
You can also see that the closest Denny's and LaQuinta is approximately 50 feet apart. 
```{r}
dennys_neighbor[,3][which(dennys_neighbor[,3] <= 5280)]
```
```
> length(dennys_neighbor[,3][which(dennys_neighbor[,3] <= 5280)])
[1] 239
> 238/nrow(dennys_neighbor)
[1] 0.1463715
```
There are only 239 Denny's that have a LaQuinta within a 1 mile radius. That's only 14.6% of all Denny's. 

```{r}
lq_neighbor[,3][which(lq_neighbor[,3] <= 5280)]
```

```
> length(lq_neighbor[,3][which(lq_neighbor[,3] <= 5280)])
[1] 238
> 238/nrow(lq_neighbor)
[1] 0.2735632
```
Similarly, there are 238 LaQuintas with a Denny's within a 1 mile radius. This is 27.4% of all LaQuintas.    

All of the information can be seen when you run the above function and view the two histograms that appear. The x-axis tells you the distance, and the y-axis tells you the frequency.   
   
As you can see, we disagree with Hedberg's claim that "La Quinta is Spanish for next to Denny's". While it may occur in a way that presents itself as a common occurrence, analysis of all locations has proven it to be not true. 

